Community-Centric Innovation

### 2014-02-18 paper.tex
about to change the paper
talking about community networks

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Bottom-up Broadband: the User-Centric Internet}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

\begin{abstract}
%\boldmath
A new generation of networks is flourishing: Bottom-up networks.
These networks take different forms, grow in different places and use different technologies.
Still, they have something in common which is that they have their origins in the users.

The users own and manage the network following community practices.
This particularity confers bottom-up networks unique properties.
For example, they can grow in underserved areas, offer an additional layer of resilience against privacy invading practices and circumvent censorship.

\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{T}{he} sustained growth of community networks represent an interesting phenomena in the Internet ecosystem.
They have their origins in reduced groups of experts and enthusiasts that built them as an experiment or for self-provisioning Internet access.
Over the years, they have proved their viability and validity as a complement to existing commercial Internet offering.

The fact that independent and unrelated initiatives have appeared around the globe points to the fact that their existence is just a natural part of the Internet.

\section{From consumer to participant}

\section{Bridging the digital divide}

\section{Resilience against privacy invading practices and censorship}

\section{The educational value of bottom-up networks}

\section{Global Co-ordination}

\section{The Bottom-up Broadband initiative in Europe}

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Community-Centric Innovation}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

\begin{abstract}
%\boldmath
The Internet is profoundly transforming the economic and social structures.
The ability to instantaneously share data across the globe enables new ways of collaboration and production models.
The individuals can join loosely organized working groups and communities and contribute to the evolution and progress of the humankind.
By choosing which communities they join and which communities they contribute to, the users can also be participants and producers.
Furthermore, this model enables a new way of democracy in which the people vote not only with their ballots, but also with their daily jobs and contributions.
In this work we describe examples of bottom-up initiatives in which the users play a central role in the development of the project.
These are initiatives by and for the people.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{T}{he} growth of the Internet has dramatically accelerated the path of technological evolution.
The amount of information and knowledge available on the Net is increasing as well as the number of people that can access that knowledge.
Still, it is important to realize that the Internet is not simply an incremental evolution that brings faster progress.
The Internet is a disruptive transformation that is completely transforming the way we organize ourselves and relate to each other.

The fact that the information can be shared with little restrictions in terms of space, time and cost, makes it possible for people to organize in horizontal structures.
In this horizontal models, also known as peer-to-peer, all the participants can access to the information required to contribute to the decision-making process.
The heterogeneous background of the participants makes it possible to consider a wider space of solutions and the extensive scrutiny reduces the number of mistakes.

In this work we consider several examples in which the users organize in communities to construct user-centric networks and networked services.

\section{Community Networks}
Probably the best example of user-centric networking are community networks.
The community networks are deployed by the same users that use them.
Consequently, they are carefully tailored to satisfy the user needs.
This needs are not necessarily only about connectivity.
The users might be interested in learning about the network, and in researching  new services, applications and protocols.

Community wireless are particularly useful in underserved areas with little presence of commercial operators.
In rural, distant or poor areas community networks can be the only viable alternative to access the Internet.
In this places, the inhabitants have a strong incentive in spending time and energies in acquiring the necessary skills required to deploy and operate a network.

Another aspect that can draw the attention to community networks is privacy.
As commercial networks are routinely involved in mass surveillance programs, the citizens see in community networks a way to opt out from privacy-invading campaigns.
Finally, community networks are also means of getting around censorship.
Totalitarian regimes include censorship into the networks to control which information is distributed among the population.
Filtering the information and ordering Internet blackouts have been used as means to hinder social movements and placate population unrest.
